---
title: 'Introduction to <img src = "rstudio_logo.png" alt = "R" width = "40"> Programming'
author: "PSE - M1 APE"
date: "Louis Sirugue"
params:
  dark: false
  update_functions: false
output: 
  html_document:
    theme: cosmo
    css: ../source/style.css
    includes:
      in_header: ../source/header.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
lapply(c("tidyverse", "kableExtra"), library, character.only = T)
root <- "https://louissirugue.github.io/intro_to_R/"
```

```{css, echo = F, eval = !params$dark}
.share-box {
  width: 40px;
  height: 40px;
  background: #DFE6EB;
  border-radius: 80px;
  text-align:left;
  padding:24px;
  box-shadow: 4px 4px 6px #c7c4c4, 
              -4px -4px 6px #E7EDF0;
}

.share-box:active {
  box-shadow: inset 4px 4px 6px #c7c4c4, 
              inset -4px -4px 6px #E7EDF0
}
```

```{css, echo = F, eval = params$dark}
body{background-color:#201914;filter:invert(1)}

.share-box {
  width: 40px;
  height: 40px;
  background: #DFE6EB;
  border-radius: 80px;
  text-align:left;
  padding:24px;
  box-shadow: 4px 4px 6px #F3F5F7, 
              -4px -4px 6px #D8DEE3;
}

.share-box:active {
  box-shadow: inset 4px 4px 6px #F3F5F7, 
              inset -4px -4px 6px #D8DEE3
}
```

```{r, echo = F}
dark <- ifelse(params$dark, "_dark", "")
astre <- ifelse(params$dark, "sun", "moon")
```

````{=html}
```{r, echo=FALSE, results='asis'}
writeLines(paste0('<div class = "share-box" style ="margin-top:-11em;margin-left: auto; margin-right: 3em;"><a href = "https://louissirugue.github.io/intro_to_R/home', ifelse(params$dark, "", "_dark"), '.html"><p style = "margin-left:-.51cm;margin-top:-.51cm;width:40px"> <img src = "', astre, '.png"> </p> </a></div>'), 'astre.html')
xfun::file_string('astre.html')
```
````

<p style = "margin-bottom:5cm;">


<blockquote>
<i><p style="margin-left:1cm;"></p>The objective of this course is to provide you with the necessary bases in R programming to be able to carry out your empirical homeworks and research projects. The course is structured in 4 classes of 2 hours. No preliminary knowledge in programming is required, but part of the content of the last lecture is based on what you're going to see in Econometrics I. Lectures will take place in the computer room R2-07 at Campus Jourdan, but feel free to bring up your own laptop (in that case install <a href = "https://cran.r-project.org/">R</a> and <a href = "https://www.rstudio.com/products/rstudio/download/#download">R Studio</a> before the first class).</i>
</blockquote>

### Material

```{r}
s1 <- data.frame(a = c(" ", 
                       paste("Lecture", 1:4)),
                 
                 b = c(" ",
                       "Data manipulation",
                       "Data visualization",
                       "R Markdown & LaTeX",
                       "Econometrics in R"),
                 
                 c = c("Slides", 
                       paste0("[html](", root, "lecture1/slides", dark, ".html) - [pdf](", root, "lecture1/slides", dark, ".pdf)"),
                       paste0("[html](", root, "lecture2/slides", dark, ".html) - [pdf](", root, "lecture2/slides", dark, ".pdf)"),
                       paste0("[html](", root, "lecture3/slides", dark, ".html) - [pdf](", root, "lecture3/slides", dark, ".pdf)"),
                       paste0("[html](", root, "lecture4/slides", dark, ".html) - [pdf](", root, "lecture4/slides", dark, ".pdf)")),
                 
                 e = c("Data",  
                       paste0("[data](", root, "lecture1/data.zip)"),
                       paste0("[data](", root, "lecture2/data.zip)"),
                       paste0("[data](", root, "lecture3/data.zip)"),
                       paste0("[data](", root, "lecture4/data.zip)")),
                 
                 f = c("Group 1",
                       "01/09<br>09:00-11:00",
                       "02/09<br>09:00-11:00",
                       "07/09<br>14:00-16:00",
                       "07/10<br>09:00-11:00"),
                 
                 g = c("Group 2",
                       "01/09<br>11:15-13:15",
                       "02/09<br>11:15-13:15",
                       "05/09<br>09:00-11:00",
                       "07/10<br>11:15-13:15"),
                 
                 h = c("Group 3",
                       "01/09<br>14:30-16:30",
                       "02/09<br>14:30-16:30",
                       "05/09<br>16:00-18:00",
                       "07/10<br>14:30-16:30"))

kable(s1, col.names = NULL, align = "llccccc", escape = F) %>%
  row_spec(1, font_size = 17) %>%
  column_spec(1, width_max = "6em", width = "6em", width_min = "6em") %>%
  column_spec(2, width_max = "15em", width = "15em", width_min = "15em") %>%
  column_spec(3, width_max = "6em", width = "6em", width_min = "6em") %>%
  column_spec(4, width_max = "4em", width = "4em", width_min = "4em") %>%
  column_spec(5:7, width_max = "6em", width = "6em", width_min = "6em") %>%
  column_spec(1:7, extra_css = "vertical-align:middle;") %>%
  kable_styling(bootstrap_options = c("condensed", "hover"), full_width = F)
```

### Contents

<div class="row">
<div class="column" style = "margin-left:2cm;">

```{r, include = F}
roman <- c("I.", "II.")

slides <- tibble(x = NA_character_, type = NA_character_, 
                 link = NA_character_, bullet = NA_character_,
                 last = NA_character_, n = NA) %>% na.omit()

for (i in 1:length(roman)) {
  
    lecture <- tibble(x = readLines(paste0("../Lecture", i, "/slides.Rmd")))
    
    slides <- slides %>%
      bind_rows(tibble(x = paste(roman[i], gsub('title: \"|\"', "", lecture$x[2])),
                       type = "title",
                       link = paste0(root, "lecture", i, "/slides.html#1"),
                       bullet = "detail",
                       last = "",
                       n = NA)) %>%
      bind_rows(lecture %>%
                  mutate(type = case_when(grepl("^-", x) ~ "Dashed",
                                          grepl("^### [0-9]\\. ", x) ~ "Section",
                                          grepl("^#### [0-9]\\.[0-9]\\. ", x) ~ "Subsection"),
                         x = gsub(":.*$", "", x)) %>%
                  na.omit() %>%
                  group_by(x) %>%
                  filter(type == "Dashed" | row_number() == 1) %>%
                  ungroup() %>%
                  group_by(type == "Dashed") %>%
                  mutate(toc_element = ifelse(type != "Dashed", row_number(), NA)) %>%
                  ungroup() %>%
                  mutate(slide = row_number() - toc_element,
                         x = sub("#* ", "", x)) %>%
                  na.omit() %>%
                  select(x, type, slide) %>%
                  mutate(section = substr(x, 1, 1)) %>%
                  group_by(section) %>%
                  mutate(n = n(),
                         last = ifelse(row_number() == n(), "Section", "")) %>%
                  ungroup() %>%
                  mutate(bullet = ifelse(type == "Section" & n > 1, "detail", "item"),
                         link = paste0(root, "lecture", i, "/slides", 
                                       ifelse(params$dark, "_dark", ""), ".html#", slide),
                         last = ifelse(row_number() == n(), "title", last)) %>%
                  select(x, type, link, bullet, last, n))
}
slides <- slides %>% mutate(x = gsub("`", "", x))

htmlcode <- ""
for (i in 1:nrow(slides)) {
  
  # Open level 1
  if (slides$type[i] == "title") {
    if (slides$bullet[i] == "detail") {
      htmlcode <- paste0(htmlcode, '<details open style = "margin-left:1cm;"><summary><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></summary>')
    }
    if (slides$bullet[i] == "item") {
      htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.4cm;"><li>', slides$x[i], '</li></ul><p style = "margin-bottom:.75cm;"></p>')
    }
  }
  
  # Open level 2
  
  if (slides$type[i] == "Section") {
    if (slides$bullet[i] == "detail") {
      htmlcode <- paste0(htmlcode, '<details style = "margin-left:1cm;"><summary><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></summary>')
    }
    if (slides$bullet[i] == "item") {
      htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.4cm;margin-bottom:0cm;"><li><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></li></ul>')
    }
  }
  
  # Level 3
  if (slides$type[i] == "Subsection") {
    htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.6cm;margin-bottom:0cm;"><li><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></li></ul>')
  }
  
  # Close level 2
  if (slides$last[i] == "Section" & slides$n[i] > 1) {
    htmlcode <- paste0(htmlcode, '</details>')
    }
  
  # Close level 1
  if (slides$last[i] == "title") {htmlcode <- paste0(htmlcode, '</details><p style = "margin-bottom:.75cm;"></p>')}
}
writeLines(htmlcode, "toc.html")
```

````{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('toc.html')
```
````
</div>

<div class="column" style = "margin-left:-2cm;">
```{r, include = F}
roman <- c("III.", "IV.")

slides <- tibble(x = NA_character_, type = NA_character_, 
                 link = NA_character_, bullet = NA_character_,
                 last = NA_character_, n = NA) %>% na.omit()

for (i in (1:length(roman))+2) {
  
    lecture <- tibble(x = readLines(paste0("../Lecture", i, "/slides.Rmd")))
    
      fourticks <- lecture %>%
        mutate(row = row_number()) %>%
        filter(grepl("^````", x)) %>%
        mutate(x = rep(1:(nrow(.)/2), each = 2),
               group = ifelse(row_number() %% 2, "start", "stop")) %>%
        pivot_wider(names_from = "group", values_from = "row")
      
      for (j in nrow(fourticks):1) {
        lecture <- lecture %>%
          filter(!row_number() %in% fourticks$start[j]:fourticks$stop[j])
      }
    
    slides <- slides %>%
      bind_rows(tibble(x = paste(roman[i - 2], gsub('title: \"|\"', "", lecture$x[2])),
                       type = "title",
                       link = paste0(root, "lecture", i, "/slides.html#1"),
                       bullet = "detail",
                       last = "",
                       n = NA)) %>%
      bind_rows(lecture %>%
                  mutate(type = case_when(grepl("^-", x) ~ "Dashed",
                                          grepl("^### [0-9]\\. ", x) ~ "Section",
                                          grepl("^#### [0-9]\\.[0-9]\\. ", x) ~ "Subsection"),
                         x = gsub(":.*$", "", x)) %>%
                  na.omit() %>%
                  group_by(x) %>%
                  filter(type == "Dashed" | row_number() == 1) %>%
                  ungroup() %>%
                  group_by(type == "Dashed") %>%
                  mutate(toc_element = ifelse(type != "Dashed", row_number(), NA)) %>%
                  ungroup() %>%
                  mutate(slide = row_number() - toc_element,
                         x = sub("#* ", "", x)) %>%
                  na.omit() %>%
                  select(x, type, slide) %>%
                  mutate(section = substr(x, 1, 1)) %>%
                  group_by(section) %>%
                  mutate(n = n(),
                         last = ifelse(row_number() == n(), "Section", "")) %>%
                  ungroup() %>%
                  mutate(bullet = ifelse(type == "Section" & n > 1, "detail", "item"),
                         link = paste0(root, "lecture", i, "/slides", 
                                       ifelse(params$dark, "_dark", ""), ".html#", slide),
                         last = ifelse(row_number() == n(), "title", last)) %>%
                  select(x, type, link, bullet, last, n))
}
slides <- slides %>% mutate(x = gsub("`", "", x))

htmlcode <- ""
for (i in 1:nrow(slides)) {
  
  # Open level 1
  if (slides$type[i] == "title") {
    if (slides$bullet[i] == "detail") {
      htmlcode <- paste0(htmlcode, '<details open style = "margin-left:1cm;"><summary><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></summary>')
    }
    if (slides$bullet[i] == "item") {
      htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.4cm;"><li>', slides$x[i], '</li></ul><p style = "margin-bottom:.75cm;"></p>')
    }
  }
  
  # Open level 2
  
  if (slides$type[i] == "Section") {
    if (slides$bullet[i] == "detail") {
      htmlcode <- paste0(htmlcode, '<details style = "margin-left:1cm;"><summary><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></summary>')
    }
    if (slides$bullet[i] == "item") {
      htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.4cm;margin-bottom:0cm;"><li><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></li></ul>')
    }
  }
  
  # Level 3
  if (slides$type[i] == "Subsection") {
    htmlcode <- paste0(htmlcode, '<ul style = "margin-left:.6cm;margin-bottom:0cm;"><li><a href = "', 
                         slides$link[i], '">', slides$x[i], '</a></li></ul>')
  }
  
  # Close level 2
  if (slides$last[i] == "Section" & slides$n[i] > 1) {
    htmlcode <- paste0(htmlcode, '</details>')
    }
  
  # Close level 1
  if (slides$last[i] == "title") {htmlcode <- paste0(htmlcode, '</details><p style = "margin-bottom:.75cm;"></p><br>')}
}
writeLines(htmlcode, "toc2.html")
```

````{=html}
```{r, echo=FALSE, results='asis'}
xfun::file_string('toc2.html')
```
````
</div>
</div>

<p style = "margin-bottom:-1.5em;"></p>

### Cheatsheets

<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/rstudio.pdf"><img style = "margin-left:7cm;margin-top:2.3cm;" src = "../cheatsheets/rstudio.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/baser.pdf"><img style = "margin-left:-1.45cm;margin-top:-2.3cm;" src = "../cheatsheets/baser.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/readr.pdf"><img style = "margin-top:2.3cm;margin-left:-1.45cm;"  src = "../cheatsheets/readr.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/dplyr.pdf"><img style = "margin-top:-2.3cm;margin-left:-1.45cm;" src = "../cheatsheets/dplyr.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/ggplot.pdf"><img  style = "margin-top:2.3cm;margin-left:-1.45cm;"src = "../cheatsheets/ggplot.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/rmarkdown.pdf"><img style = "margin-top:-2.3cm;margin-left:-1.45cm;" src = "../cheatsheets/rmarkdown.png" width = "100"/></a>
<a href = "https://louissirugue.github.io/intro_to_R/cheatsheets/latex.pdf"><img style = "margin-top:2.3cm;margin-left:-1.45cm;" src = "../cheatsheets/latex.png" width = "100"/></a>

### Functions glossary

<p style = "margin-bottom:1cm;"></p>

```{r, echo = F, eval = params$update_functions, results = "hide"}
library(readxl)

functions <- read_xlsx("functions.xlsx")

# Get the last R documentation version of each package
packages <- functions %>% select(package) %>% unique()
version <- lapply(packages$package, function(a) {
  return(gsub('^.*versionsArray":\\["|".*$', '', 
              paste(readLines(paste0('https://www.rdocumentation.org/packages/',  a)), collapse = "")))
})

# Get the alias of each function
lapply(packages$package, require, character.only = TRUE)
alias <- lapply(functions$`function`, function(a) {
  hf <- as.character(help(a))
  keep <- grepl(paste0(functions %>% filter(`function` == a) %>% pull(package), "/help"), hf)
  return(gsub("^.*/", "", hf[keep]))
})

# Form URL
functions <- functions %>%
  left_join(packages %>% mutate(version = unlist(as.character(version)))) %>%
  mutate(alias = unlist(alias)) %>%
  # Manual modifications
  mutate(alias = ifelse(`function` %in% c("n", "log", "scale_x_discrete", 
                                          "element_rect", "element_line", "rel"), 
                        `function`, alias),
         alias = ifelse(`function` == "scale_y_discrete", "scale_x_discrete", alias),
         alias = ifelse(`function` == "scale_color_manual", "scale_colour_manual", alias),
         alias = ifelse(`function` == "scale_color_gradient", "scale_colour_gradient", alias),
         alias = ifelse(alias == "reorder.factor", "reorder.default", alias),
         version = ifelse(`function` %in% c("scale_color_manual", "geom_text", "geom_histogram"), 
                          "0.9.1", version),
         version = ifelse(`function` %in% c("element_rect", "element_line", "rel"), 
                          "2.1.0", version))%>%
  filter(`function` != "rename") %>%
  # Now form urls
  mutate(doc = paste0("https://www.rdocumentation.org/packages/", package, 
                      "/versions/", version, 
                      "/topics/", alias))

write.csv(functions, "processed_functions.csv", row.names = F)
```


```{r, echo = F}
library(DT)
datatable(read.csv("processed_functions.csv") %>% 
              mutate(doc = paste0('<a href = "', doc, '">', alias, '</a>'),
                     lecture = paste("Lecture", lecture)) %>%
              rename(`function` = function., documentation = doc) %>%
              select(lecture, `function`, package, documentation), 
          escape = F, rownames = F, options = list(pageLength = 5), 
          class = list(stripe = FALSE)) %>% 
  formatStyle(c("lecture", "function", "package", "documentation"), 
              backgroundColor = "#DFE6EB")
```

### References

This course is adapted from [the one I teach in CPES-2](https://louissirugue.github.io/metrics_on_R/home.html), inspired by:

 * [Introduction to R](https://introtor-pse.appspot.com/), <i>by H. Bull and P. Charousset.</i> Paris School of Economics (2020)
 * [Advanced Mircoeconometrics](https://enseignements.ehess.fr/2020-2021/ue/798), <i>by D. Margolis and F. Libois.</i> Paris School of Economics (2020)
 * [How to Lie with Graphics](http://data.visualisation.free.fr/Papers/Part1-HowToLie.pdf), <i>by C. Bontemps.</i> Toulouse School of Economics (2020)
 * [Introduction to Econometrics with R](https://scpoecon.github.io/ScPoEconometrics/), <i>by F. Oswald, V. Viers, P. Villedieu, and G. Kenedi.</i> SciencesPo Dep. of Economics (2020)
 * Introduction to Causal Inference, <i>by L. Zabrocki.</i> Paris Sciences et Lettres - CPES (2020)
 * Geolocalized Datasets and Applications for Economics, <i>by F. Libois and E. Madinier.</i> Paris School of Economics (2020)
 * Countless posts on [stackoverflow](https://stackoverflow.com/) and [R-bloggers](https://www.r-bloggers.com/)
 
<br><br><br><br>